Young children rationally integrate semantic and pragmatic information sources to infer the meaning of words

### 2020-08-26 paper/manuscript.Rmd
first draft reformatted paper

When researchers do look at information integration in children [@ganea2007infants; @graham2017words; @grosse201021; @jara2020social; @khu2020preschoolers; @matthews2006effect; @nilsen2009preschoolers], they mostly focus on how children interpret an ambiguous referring expression in light of social-contextual information. In one of these studies [@nadig2002evidence], children faced a 2 x 2 display with a ball, a pen and two glasses in it. The speaker, sitting on the opposite side of the display, saw only 3 of the compartments: the ball, the pen, and one of the glasses. When the speaker asked for "the glass", children had to integrate the semantics of the utterance with the perspective of the speaker to correctly infer which of the glasses - from their perspective - the speaker was referring to. This study, along with others, found that preschoolers use both information sources. What these studies do not do, however, is specify the process by which children integrate the different information sources. When discussing their results, authors refer to social-pragmatic theories of language use and learning [@grice1991studies; @sperber2001relevance; @clark1996using; @clark2009first; @tomasello2009constructing], all of which assume that information is integrated as part of a social inference process, but none of which clearly defines the process. As a consequence, we have no explicit and quantitative theory of how different information sources (and word learning mechanisms) are integrated.

@@

Our *rational integration model* arbitrates between information sources via Bayesian inference (see Fig. 1F for model formula). A listener ($L_{1}$) reasons about the referent of a speaker's ($S_{1}$) utterance. This reasoning is contextualized by the prior probability of each referent $\rho$ which we take to be a function of the common ground shared between the listener and the speaker. To decide between referents, the listener ($L_{1}$) reasons about what a rational speaker ($S_{1}$) with informativeness $\alpha$ would say given an intended referent. This speaker is assumed to compute the informativity for each available utterance and then choose the most informative one. The informativity of each utterance is given by imagining which referent a listener, who interprets words according to their literal semantics (what we call a literal listener, $L_{0}$), would infer upon hearing the utterance. Naturally, this reasoning depends on what kind of semantic knowledge (for object j) $\theta$~j~ the speaker ascribes to the (literal) listener. Taken together, this model provides a quantitative theory of information integration during language learning. Furthermore, the model presents an explicit and substantive theory of development. It assumes that, while children’s sensitivity to the individual information sources increases with age, the way integration proceeds remains constant [@bohn2019pervasive; @bohn_tessler_merrick_frank_2019]. In the model, this is accomplished by making the parameters capturing sensitivity to speaker informativeness ($\alpha$~i~, Fig. 1D), the common ground ($\rho$~i~, Fig. 1C), and object specific semantic knowledge ($\theta$~i,j~, Fig. 1E) a function of the child’s age.

```{r fig1, include = T, fig.align = "center", fig.cap = "Experimental task and model. (A and B) Screenshots from the experimental task. (i) The speaker encounters one object and then leaves the scene (ii). While the speaker is away, a second object appears (iii), when returning, the speaker uses a novel word to request an object (iv). Sections (i) to (iii) establish common ground between the speaker and the listener, in that one object is new in context (green). The request in (iv) licenses an inference based on expectations about how informative speakers are (red). Listeners' semantic knowledge enters the task because the identity of the  known object on one of the tables is varied from well-known objects like a duck to relatively unfamiliar objects like a chess pawn (total of 12 objects - blue). (A) shows the condition in which common ground information is congruent (i.e. point to the same object) with speaker informativeness and (B) shows the incongruent condition. The congruent and incongruent conditions are each paired with the 12 known objects, resulting in 24 unique conditions. (C to E) show developmental trajectories for each information source, estimated from separate experiments (see supplement). (F) gives the model equation for the rational integration model and links information sources to model parameters.", out.width="100%"}
knitr::include_graphics("./figures/figure1.png")
```

To generate data to test the predictive and explanatory power of our model, we developed a world-learning paradigm in which we jointly manipulated the three information sources (Fig. 1). Children interacted on a tablet computer with a series of storybook speakers [@frank2016using]. The speakers used novel words (e.g., "wug") in the context of two potential referents. This situation is depicted in Fig. 1A-iv, in which a speaker (here, a frog) appears with a known object (a duck, left) and an unfamiliar object (the diamond-shaped object, right). To identify the referent, the child could rely upon the three information sources described above. First, they may infer that a cooperative speaker would have used the word duck to refer to the known object (the duck); the fact that the speaker did not say duck then suggests that the speaker is most likely referring to a different object (the unfamiliar object). This basic inference is oftentimes referred to as a mutual exclusivity inference [@markman1988children; @halberda2003development]. More generally speaking, the listener can infer the referent of the novel word by assuming that the speaker is a rational agent who chooses the most informative words to refer to objects [@clark1987principle; @frank2014inferring]. Second, the child may rely upon what has already been established in the common ground with the speaker. Listeners expect speakers to communicate about things that are new to the common ground [@akhtar1996role; @diesendruck2004two]. Thus, the inference about the novel word referring to the unfamiliar object also depends on which object is new in context (Fig. 1A and B i-iii). Finally, the child may rely upon their previously acquired semantic knowledge, that is, how sure they are that the known object is called "duck". If the known object is something less familiar, such as a chess piece (e.g., a pawn), a 3-year-old child may draw a weaker inference, if they draw any inference at all [@lewis2020-me]. Taken together, the child has the opportunity to integrate their assumptions about (1) informative communication, (2) their understanding of the common ground, and (3) their existing semantic knowledge. In some versions of the experiment, some of these information sources were aligned (Fig. 1A) but in some they were in conflict (Fig. 1B).

@@

We tested the model in its ability to predict children’s actual behavior - at an age (between 2 and 5 years) at which they learn language at an exponential rate [@bloom2002children]. We estimated children’s (N=148) developing sensitivity to individual information sources in separate experiments (see Fig. 1C-E). We then generated parameter-free, quantitative, *a priori* model predictions (developmental trajectories) representing the model's expectations about how children should behave in a new situation in which all three information sources had to be integrated. We generated model predictions for 24 experimental conditions: 12 objects of different familiarities (requiring different levels of semantic knowledge), with novelty either conflicting or coinciding; Fig. 1. We compared these predictions to newly collected data from N = 220 children from the same age range. The results showed a very close alignment between model predictions and the data across the entire age range. That is, the average developmental trajectories predicted by the model resembled the trajectories found in the data (Fig. S6). For a more quantitative analysis, we binned predictions and data by year and correlated them. We found a high correlation showing that the model explained 79% of the variance in the data (Fig. 2). These results seem to support the assumptions in the model that children integrate all available information sources. However, it is still possible that simpler models might make equally good or better predictions.

Thus, we formalized the alternative view that children selectively ignore information sources in the form of three lesioned models (Fig. 2). These models assume that children follow the heuristic “ignore X” (with X being one of the information sources) when multiple information sources are presented together.

The marginally less-complex *no word knowledge model* uses the same model architecture as the rational integration model. It takes in expectations about speaker informativeness and common ground, but omits semantic knowledge that is specific to the familiar objects (i.e., uses only general semantic knowledge). The object specific parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect) and then there is object-specific variation around this trajectory (random effects). The no word knowledge model takes in the overall trajectory for semantic knowledge but ignores object-specific variation. That is, the model assumes a listener whose inference does not vary depending on the particular familiar object but only depends on the age specific average semantic knowledge.The *no common ground model* takes in object specific semantic knowledge and speaker informativeness but ignores common ground information. Instead of assuming that one object has a higher prior probability to be the referent because it is new in context, the speaker thinks that both objects are equally likely to be the referent. As a consequence, the listener does not differentiate between situations in which common ground is aligned or in conflict with the other information sources. Finally, according to the *no speaker inforamtiveness model*, the listener does not assume that the speaker is communicating in an informative way and hence ignores the information provided by the utterance. As a consequence, the inference is solely based on common ground expectations.

@@

(ref:figlab2) **Predicting information integration**. Correlation between model predictions and child inference data for all 24 conditions and for each age group (binned by year) for the rational integration model and the three lesioned models. Horizontal and vertical error bars show 95% HDI. Inset shows an example of model predictions as developmental trajectories (see Fig. 3). The no word knowledge model includes developmental change for overall vocabulary development but ignores the identity of the known object. The no common ground model assumes that children ignore information from common ground and the no speaker informativeness model assumes that children ignore inferences that are licensed by expectations about informative communication. BF~10~ gives the Bayes Factor in favor of the integration model based on the marginal likelihood of the data under each model.

@@

The biased integration model makes reasonable predictions that explain a large proportion of the variance in the data (Fig. 3B). The parameter $\phi$ - indicating the bias to one of the inferences - was estimated to favor the mutual exclusivity inference (mode = `r phi%>%pull(mode)`; 95% HDI: `r phi%>%pull(lci)` - `r phi%>%pull(uci)`, see Fig. 3B). However, the rational integration model presented a much better fit to the data, both in terms of correlation and marginal likelihood of the data (Fig. 3B). Thus, a fully integrated - as opposed to a segmented, biased - integration process explains the data better.

@@

The developmental bias model explains a substantial portion of the variance in the data (Fig. 3B). The estimated developmental trajectory for the bias parameter $\phi$ (Fig. 3B) suggests that younger children put a higher emphasis on common ground information, while older children rely more on the mutual exclusivity inference. The relative importance of the two inferences seems to switch at around 3 years. Yet again, when we directly compare the competitor models, we find that the data is several orders of magnitude more likely under the rational integration model (Fig. 3B). Interestingly, when we compared the rational integration model with its parameters not constrained by the experimental data (i.e. the "pure" prediction model described above), it also outperformed both bias models (see Fig. S10 and Tab. S6). Taken together, these results show that the integration process is best described as fully integrated and stable over time.

(ref:figlab3) **Explaining information integration across development**. (A) Model predictions from the rational integration model (colored lines) next to the behavioral data (dotted black lines with 95% CI in gray) for all 24 experimental conditions. Top row (blue) shows conditions in which common ground information is congruent with speaker informativeness, bottom row (red) shows conditions in which they are incongruent. Known objects are ordered based on their rated age of acquisition (left to right). Light dots represent individual data points. (B) Correlations between model predictions binned by age and condition for the integration model and (C) the two mixture models. Vertical and horizontal error bars show 95% HDIs. BF~10~ gives the Bayes Factor in favor of the rational integration model based on the marginal likelihood of the data under each model. (D) Posterior distribution of the bias parameter in the biased integration model and developmental trajectories for the bias parameter in the developmental bias model based on 300 random draws from the posterior distribution of the intercept and slope.

@@

All experiments were presented as an interactive picture book on a tablet computer [@frank2016using]. Fig. 1A and B show screenshots from the actual experiments. The general setup involved an animal standing on a little hill between two tables. For each animal character, we recorded a set of utterances (one native English speaker per animal) that were used to talk to the child and make requests. Each experiment started with two training trials in which the speaker requested known objects (car and ball).

@@

Experiment 1 tested the so-called mutual exclusivity inference [@markman1988children; @lewis2020-me]. On one table, there was a familiar object, on the other table, there was a novel object (drawn for the purpose of the study) (Fig. 1A/B iv and Fig. S1A). The speaker requested an object by saying "Oh cool, there is a [non-word] on the table, how neat, can you give me the [non-word]?". Children responded by touching one of the objects. The location of the novel object (left or right table) and the animal character were counterbalanced. We coded as correct choice if children chose the novel object as the referent of the novel word. Each child completed 12 trials, each with a different familiar and a different novel object. Familiar objects were selected to vary along the dimension of how likely children were to know the word for it. This including objects that most 2-year-olds can name (e.g. a duck) as well as objects that only very few 5-year-olds can name (e.g. a pawn). The selection was based on age of acquisition ratings from Kuperman and colleagues [-@kuperman2012age]. While these ratings do not capture the absolute age when children acquire these words, they capture the relative order in which words are learned. Fig. S2A in the supplementary material shows the words and objects used in the experiment. We induced this variation to estimate the role of semantic knowledge in a mutual exclusivity inference.

Experiment 2 tested children's sensitivity to common ground that is build up over the course of a conversation. In particular, we tested whether children keep track of which object is new to a speaker and which they have encountered previously [@akhtar1996role; @diesendruck2004two]. The general setup was the same as in Experiment 1 (Fig. S1B). The speaker was positioned between the tables. There was a novel object (drawn for the purpose of the study) on one of the tables while the other table was empty. Next, the speaker turned to one of the tables and either commented on the presence ("Aha, look at that.") or the absence ("Hm, nothing there") of an object. Then the speaker disappeared. While the speaker was away, a second novel object appeared on the previously empty table. Then the speaker returned and requested an object in the same way as in Experiment 1. The positioning of the novel object in the beginning of the experiment, the speaker as well as the location the speaker turned to first was counterbalanced. Children received five trials, each with a different pair of novel objects. We coded as correct choice if children chose the object that was new to the speaker as the referent of the novel word.

Experiment 3 combined the procedures from Experiment 1 and 2. It followed the same procedure as Experiment 2 but involved the same objects as Experiment 1 (Fig. 1 i-iv and Fig. S1C). In the beginning, one table was empty while there was an object (novel or familiar) on the other one. After commenting on the presence or absence of an object on each table, the speaker disappeared and a second object appeared (familiar or novel). Next, the speaker re-appeared and made the usual request. In the congruent condition, the familiar object was present in the beginning and the novel object appeared while the speaker was away (Fig. 1A and Fig. S1C - left). In this case, both the mutual exclusivity and the common ground inference pointed to the novel object as the referent. In the incongruent condition, the novel object was present in the beginning and the familiar object appeared later. In this case, the two inferences pointed to different objects (Fig. 1B and Fig. S1C - right). This resulted in a total of 2 (congruent vs incongurent) x 12 (familiar objects) = 24 different conditions. Participants received up to 12 test trials, six in each condition, each with a different familiar and novel object. Familiar objects were the same as in Experiment 1. The positioning of the objects on the tables, the speaker and the location the speaker first turned to were counterbalanced. Participants could stop the experiment after six trials (three per condition). If a participant stopped after half of the trials, we tested an additional participant to reach a pre-registered number of data points per age group (2-, 3- and 4-year-olds).

@@

As a first step, we used the data from Experiment 1 and 2 to estimate children's developing sensitivity to each information source. To estimate the parameters for semantic knowledge ($\theta$) and speaker informativeness ($\alpha$), we adapted the rational integration model to a situation in which both objects (novel and familiar) have equal prior probability (i.e., no common ground information). We used the data from Experiment 1 to then infer the parameters. That is, we inferred the intercepts and slopes for speaker informativeness (linear regression) and semantic knowledge (logistic regression) that generated model predictions to match the responses generated in Experiment 1. To estimate the parameters representing sensitivity to common ground ($\rho$), we used a simple logistic regression to infer which combination of intercept and slope would generate predictions that corresponded to the average proportion of correct responses measured in Experiment 2. A more detailed description of these parameters were estimated (including a graphical model) can be found in the supplementary material.

To generate model predictions for the rational integration model we combined the parameters according to the model formula (Fig. 1F). As mentioned above, common ground information could either be aligned or in conflict with the other information sources. In the congruent condition, the novel object was also the discourse novel object and thus had the prior probability $\rho$. In the incongruent condition, the novel object was the "old" object and thus had the prior probability of $1 - \rho$.

>>>>

When information integration in children is studied directly [@ganea2007infants; @graham2017words; @grosse201021; @jara2020social; @khu2020preschoolers; @matthews2006effect; @nilsen2009preschoolers], the focus is mostly on how children interpret an ambiguous referring expression in light of social-contextual information. In one of these studies [@nadig2002evidence], children faced a 2 x 2 display with a ball, a pen and two glasses in it. The speaker, sitting on the opposite side of the display, saw only 3 of the compartments: the ball, the pen, and one of the glasses. When the speaker asked for "the glass", children had to integrate the semantics of the utterance with the perspective of the speaker to correctly infer which of the glasses - from their perspective - the speaker was referring to. This study, along with others, found that preschoolers use both information sources. What these studies do not do, however, is specify the process by which children integrate the different information sources. When discussing their results, authors refer to social-pragmatic theories of language use and learning [@grice1991studies; @sperber2001relevance; @clark1996using; @clark2009first; @tomasello2009constructing], all of which assume that information is integrated as part of a social inference process, but none of which clearly defines the process. As a consequence, we have no explicit and quantitative theory of how different information sources (and word learning mechanisms) are integrated.

@@

Our *rational integration model* arbitrates between information sources via Bayesian inference (see Fig. 1f for model formula). A listener ($L_{1}$) reasons about the referent of a speaker's ($S_{1}$) utterance. This reasoning is contextualized by the prior probability of each referent $\rho$ which we take to be a function of the common ground shared between the listener and the speaker. To decide between referents, the listener ($L_{1}$) reasons about what a rational speaker ($S_{1}$) with informativeness $\alpha$ would say given an intended referent. This speaker is assumed to compute the informativity for each available utterance and then choose the most informative one. The informativity of each utterance is given by imagining which referent a listener, who interprets words according to their literal semantics (what we call a literal listener, $L_{0}$), would infer upon hearing the utterance. Naturally, this reasoning depends on what kind of semantic knowledge (for object j) $\theta$~j~ the speaker ascribes to the (literal) listener. Taken together, this model provides a quantitative theory of information integration during language learning. Furthermore, the model presents an explicit and substantive theory of development. It assumes that, while children’s sensitivity to the individual information sources increases with age, the way integration proceeds remains constant [@bohn2019pervasive; @bohn_tessler_merrick_frank_2019]. In the model, this is accomplished by making the parameters capturing sensitivity to speaker informativeness ($\alpha$~i~, Fig. 1d), the common ground ($\rho$~i~, Fig. 1c), and object specific semantic knowledge ($\theta$~i,j~, Fig. 1e) a function of the child’s age.

```{r fig1, include = T, fig.align = "center", fig.cap = "Experimental task and model. (a and b) Screenshots from the experimental task. (i) The speaker encounters one object and then leaves the scene (ii). While the speaker is away, a second object appears (iii), when returning, the speaker uses a novel word to request an object (iv). Sections (i) to (iii) establish common ground between the speaker and the listener, in that one object is new in context (green). The request in (iv) licenses an inference based on expectations about how informative speakers are (red). Listeners' semantic knowledge enters the task because the identity of the  known object on one of the tables is varied from well-known objects like a duck to relatively unfamiliar objects like a chess pawn (total of 12 objects - blue). (a) shows the condition in which common ground information is congruent (i.e. point to the same object) with speaker informativeness and (b) shows the incongruent condition. The congruent and incongruent conditions are each paired with the 12 known objects, resulting in 24 unique conditions. (c to e) show developmental trajectories for each information source, estimated from separate experiments (see supplement). (f) gives the model equation for the rational integration model and links information sources to model parameters.", out.width="100%"}
knitr::include_graphics("./figures/figure1.png")
```

To generate data to test the predictive and explanatory power of our model, we developed a world-learning paradigm in which we jointly manipulated the three information sources (Fig. 1). Children interacted on a tablet computer with a series of storybook speakers [@frank2016using]. The speakers used novel words (e.g., "wug") in the context of two potential referents. This situation is depicted in Fig. 1a iv, in which a speaker (here, a frog) appears with a known object (a duck, left) and an unfamiliar object (the diamond-shaped object, right). To identify the referent, the child could rely upon the three information sources described above. First, they may infer that a cooperative speaker would have used the word duck to refer to the known object (the duck); the fact that the speaker did not say duck then suggests that the speaker is most likely referring to a different object (the unfamiliar object). This basic inference is oftentimes referred to as a mutual exclusivity inference [@markman1988children; @halberda2003development]. More generally speaking, the listener can infer the referent of the novel word by assuming that the speaker is a rational agent who chooses the most informative words to refer to objects [@clark1987principle; @frank2014inferring]. Second, the child may rely upon what has already been established in the common ground with the speaker. Listeners expect speakers to communicate about things that are new to the common ground [@akhtar1996role; @diesendruck2004two]. Thus, the inference about the novel word referring to the unfamiliar object also depends on which object is new in context (Fig. 1a and b i-iii). Finally, the child may rely upon their previously acquired semantic knowledge, that is, how sure they are that the known object is called "duck". If the known object is something less familiar, such as a chess piece (e.g., a pawn), a 3-year-old child may draw a weaker inference, if they draw any inference at all [@lewis2020-me]. Taken together, the child has the opportunity to integrate their assumptions about (1) informative communication, (2) their understanding of the common ground, and (3) their existing semantic knowledge. In some versions of the experiment, some of these information sources were aligned (Fig. 1a) but in some they were in conflict (Fig. 1b).

@@

We tested the model in its ability to predict children’s actual behavior - at an age (between 2 and 5 years) at which they learn language at an exponential rate [@bloom2002children]. We estimated children’s (N=148) developing sensitivity to individual information sources in separate experiments (see Fig. 1c-e). We then generated parameter-free, quantitative, *a priori* model predictions (developmental trajectories) representing the model's expectations about how children should behave in a new situation in which all three information sources had to be integrated. We generated model predictions for 24 experimental conditions: 12 objects of different familiarities (requiring different levels of semantic knowledge), with novelty either conflicting or coinciding; Fig. 1. We compared these predictions to newly collected data from N = 220 children from the same age range. The results showed a very close alignment between model predictions and the data across the entire age range. That is, the average developmental trajectories predicted by the model resembled the trajectories found in the data (Fig. S6). For a more quantitative analysis, we binned predictions and data by year and correlated them. We found a high correlation showing that the model explained 79% of the variance in the data (Fig. 2a). These results seem to support the assumptions in the model that children integrate all available information sources. However, it is still possible that simpler models might make equally good or better predictions.

Thus, we formalized the alternative view that children selectively ignore information sources in the form of three lesioned models (Fig. 2b). These models assume that children follow the heuristic “ignore X” (with X being one of the information sources) when multiple information sources are presented together.

The marginally less-complex *no word knowledge model* uses the same model architecture as the rational integration model. It takes in expectations about speaker informativeness and common ground, but omits semantic knowledge that is specific to the familiar objects (i.e., uses only general semantic knowledge). That is, the model assumes a listener whose inference does not vary depending on the particular familiar object but only depends on the age specific average semantic knowledge.The *no common ground model* takes in object specific semantic knowledge and speaker informativeness but ignores common ground information. Instead of assuming that one object has a higher prior probability to be the referent because it is new in context, the speaker thinks that both objects are equally likely to be the referent. As a consequence, the listener does not differentiate between situations in which common ground is aligned or in conflict with the other information sources. Finally, according to the *no speaker inforamtiveness model*, the listener does not assume that the speaker is communicating in an informative way and hence ignores the information provided by the utterance. As a consequence, the inference is solely based on common ground expectations.

@@

(ref:figlab2) **Predicting information integration**. Correlation between model predictions and child inference data for all 24 conditions and for each age group (binned by year) for the rational integration model (a) and the three lesioned models (b). Horizontal and vertical error bars show 95% HDI. Inset shows an example of model predictions as developmental trajectories (see Fig. 3). The no word knowledge model includes developmental change for overall vocabulary development but ignores the identity of the known object. The no common ground model assumes that children ignore information from common ground and the no speaker informativeness model assumes that children ignore inferences that are licensed by expectations about informative communication. BF~10~ gives the Bayes Factor in favor of the integration model based on the marginal likelihood of the data under each model.

@@

The biased integration model makes reasonable predictions that explain a large proportion of the variance in the data (Fig. 3b). The parameter $\phi$ - indicating the bias to one of the inferences - was estimated to favor the mutual exclusivity inference (mode = `r phi%>%pull(mode)`; 95% highest density interval (HDI): `r phi%>%pull(lci)` - `r phi%>%pull(uci)`, see Fig. 3d). However, the rational integration model presented a much better fit to the data, both in terms of correlation and marginal likelihood of the data (Fig. 3). Thus, a fully integrated - as opposed to a segmented and biased - integration process explained the data better.

@@

The developmental bias model explains a substantial portion of the variance in the data (Fig. 3c). The estimated developmental trajectory for the bias parameter $\phi$ suggests that younger children put a higher emphasis on common ground information, while older children rely more on the mutual exclusivity inference (Fig. 3d). The relative importance of the two inferences seems to switch at around 3 years. Yet again, when we directly compare the competitor models, we find that the data is several orders of magnitude more likely under the rational integration model (Fig. 3). Interestingly, when we compared the rational integration model with its parameters not constrained by the experimental data (i.e. the "pure" prediction model described above), it also outperformed both bias models (see Fig. S10 and Tab. S6). Taken together, these results show that the integration process is best described as fully integrated and stable over time.

(ref:figlab3) **Explaining information integration across development**. (a) Model predictions from the rational integration model (colored lines) next to the behavioral data (dotted black lines with 95% CI in gray) for all 24 experimental conditions. Top row (blue) shows conditions in which common ground information is congruent with speaker informativeness, bottom row (red) shows conditions in which they are incongruent. Known objects are ordered based on their rated age of acquisition (left to right). Light dots represent individual data points. (b) Correlations between model predictions binned by age and condition for the integration model and (c) the two mixture models. Vertical and horizontal error bars show 95% HDIs. BF~10~ gives the Bayes Factor in favor of the rational integration model based on the marginal likelihood of the data under each model. (d) Posterior distribution of the bias parameter in the biased integration model and developmental trajectories for the bias parameter in the developmental bias model based on 300 random draws from the posterior distribution of the intercept and slope.

@@

All experiments were presented as an interactive picture book on a tablet computer [@frank2016using]. Fig. 1a and b show screenshots from the actual experiments. The general setup involved an animal standing on a little hill between two tables. For each animal character, we recorded a set of utterances (one native English speaker per animal) that were used to talk to the child and make requests. Each experiment started with two training trials in which the speaker requested known objects (car and ball).

@@

Experiment 1 tested the so-called mutual exclusivity inference [@markman1988children; @lewis2020-me]. On one table, there was a familiar object, on the other table, there was a novel object (drawn for the purpose of the study) (Fig. 1a/b iv and Fig. S1a). The speaker requested an object by saying "Oh cool, there is a [non-word] on the table, how neat, can you give me the [non-word]?". Children responded by touching one of the objects. The location of the novel object (left or right table) and the animal character were counterbalanced. We coded as correct choice if children chose the novel object as the referent of the novel word. Each child completed 12 trials, each with a different familiar and a different novel object. Familiar objects were selected to vary along the dimension of how likely children were to know the word for it. This including objects that most 2-year-olds can name (e.g. a duck) as well as objects that only very few 5-year-olds can name (e.g. a pawn). The selection was based on age of acquisition ratings from Kuperman and colleagues [-@kuperman2012age]. While these ratings do not capture the absolute age when children acquire these words, they capture the relative order in which words are learned. Fig. S2A in the supplementary material shows the words and objects used in the experiment. We induced this variation to estimate the role of semantic knowledge in a mutual exclusivity inference.

Experiment 2 tested children's sensitivity to common ground that is build up over the course of a conversation. In particular, we tested whether children keep track of which object is new to a speaker and which they have encountered previously [@akhtar1996role; @diesendruck2004two]. The general setup was the same as in Experiment 1 (Fig. S1b). The speaker was positioned between the tables. There was a novel object (drawn for the purpose of the study) on one of the tables while the other table was empty. Next, the speaker turned to one of the tables and either commented on the presence ("Aha, look at that.") or the absence ("Hm, nothing there") of an object. Then the speaker disappeared. While the speaker was away, a second novel object appeared on the previously empty table. Then the speaker returned and requested an object in the same way as in Experiment 1. The positioning of the novel object in the beginning of the experiment, the speaker as well as the location the speaker turned to first was counterbalanced. Children received five trials, each with a different pair of novel objects. We coded as correct choice if children chose the object that was new to the speaker as the referent of the novel word.

Experiment 3 combined the procedures from Experiment 1 and 2. It followed the same procedure as Experiment 2 but involved the same objects as Experiment 1 (Fig. 1 i-iv and Fig. S1c). In the beginning, one table was empty while there was an object (novel or familiar) on the other one. After commenting on the presence or absence of an object on each table, the speaker disappeared and a second object appeared (familiar or novel). Next, the speaker re-appeared and made the usual request. In the congruent condition, the familiar object was present in the beginning and the novel object appeared while the speaker was away (Fig. 1a and Fig. S1c - left). In this case, both the mutual exclusivity and the common ground inference pointed to the novel object as the referent. In the incongruent condition, the novel object was present in the beginning and the familiar object appeared later. In this case, the two inferences pointed to different objects (Fig. 1B and Fig. S1c - right). This resulted in a total of 2 (congruent vs incongurent) x 12 (familiar objects) = 24 different conditions. Participants received up to 12 test trials, six in each condition, each with a different familiar and novel object. Familiar objects were the same as in Experiment 1. The positioning of the objects on the tables, the speaker and the location the speaker first turned to were counterbalanced. Participants could stop the experiment after six trials (three per condition). If a participant stopped after half of the trials, we tested an additional participant to reach a pre-registered number of data points per age group (2-, 3- and 4-year-olds).

@@

As a first step, we used the data from Experiment 1 and 2 to estimate children's developing sensitivity to each information source. To estimate the parameters for semantic knowledge ($\theta$) and speaker informativeness ($\alpha$), we adapted the rational integration model to a situation in which both objects (novel and familiar) have equal prior probability (i.e., no common ground information). We used the data from Experiment 1 to then infer the parameters. That is, we inferred the intercepts and slopes for speaker informativeness (linear regression) and semantic knowledge (logistic regression) that generated model predictions to match the responses generated in Experiment 1. To estimate the parameters representing sensitivity to common ground ($\rho$), we used a simple logistic regression to infer which combination of intercept and slope would generate predictions that corresponded to the average proportion of correct responses measured in Experiment 2. For the explanation models, these parameters were additionally constrained by the data from Experiment 3. A more detailed description of these parameters were estimated (including a graphical model) can be found in the supplementary material.

To generate model predictions, we combined the parameters according to the respective model formula. As mentioned above, common ground information could either be aligned or in conflict with the other information sources. In the congruent condition, the novel object was also the discourse novel object and thus had the prior probability $\rho$. In the incongruent condition, the novel object was the "old" object and thus had the prior probability of $1 - \rho$.

The rational integration model is defined as $P^{int}_{L_1}(r \mid u; \{\rho_i, \alpha_i\, \theta_{ij}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{ij}\}) \cdot P(r \mid \rho_i)$. The three lesioned models that were used to compare how well the model predicts new data are reduced versions of this model. The no word knowledge model uses the same model architecture: $P^{no\_wk}_{L_1}(r \mid u; \{\rho_i, \alpha_i\, \theta_{i}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{i}\}) \cdot P(r \mid \rho_i)$ and the only difference lies in the parameter $\theta$. The object specific parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect - $\theta_{i}$) and then there is object-specific variation around this trajectory (random effects -- $\theta_{ij}$). Thus, the no word knowledge model takes in the overall trajectory for semantic knowledge ($\theta_{i}$) but ignores object-specific variation. The no common ground model ignores common ground information (represented by $\rho$) and is thus defined as $P^{no\_cg}_{L_1}(r \mid u; \{\alpha_i\, \theta_{ij}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{ij}\})$. For the no speaker informativeness model, the parameter $\alpha = 0$. As a consequence, the likelihood term in the model is 1 and the model therefore reduces to $P^{no\_si}_{L_1}(r \mid u; \{\rho_i\})\propto P(r \mid \rho_i)$.

As noted above, the explanation models used parameters that were additionally constrained by the data from Experiment 3. This had no effect on the way these parameters were combined in the rational integration model. The biased integration model is defined as $P^{biased}_{L_1}(r \mid u; \{\phi, \rho_i, \alpha_i, \theta_{ij} \}) = \phi \cdot P_{ME}(r \mid u; \{\alpha_i, \theta_{ij}\})  + (1-\phi) \cdot P(r \mid \rho_i)$ with $P_{ME}$ representing a mutual exclusivity inference which takes in speaker inforamtiveness and object specific semantic knowledge. This inference is then weighted by the parameter $\phi$ and added to the respective prior probability, which is weighted by $1-\phi_i$. Thus, $\phi$ represents the bias in favor of the mutual exclusivity inference. In the developmental bias model the parameter $\phi$ is made to change with age ($\phi_i$) and the model is thus defined as $P^{dev\_bias}_{L_1}(r \mid u; \{\phi_i, \rho_i, \alpha_i, \theta_{ij} \}) = \phi_i \cdot P_{ME}(r \mid u; \{\alpha_i, \theta_{ij}\})  + (1-\phi_i)$.

We compared models in two ways. First, we used Pearson correlations between model predictions and the data. For this analysis, we binned the model predictions and the data by age in years and by the type of familiar object (see Fig. 2 and 3 as well as S7 and S10). Second, we compared models based on the marginal likelihood of the data under each model -- the likelihood of the data averaging over ("marginalizing over") the prior distribution on parameters; the pairwise ratio of marginal likelihoods for two models is known as the Bayes Factor. It is interpreted as how many times more likely the data is under one model compared to the other. Bayes Factors quantify the quality of predictions of a model, averaging over the possible values of the parameters of the models (weighted by the prior probabilities of those parameter values); by averaging over the prior distribution on parameters, Bayes Factor implicitly take into account model complexity because models with more parameters will tend to have a broader prior distribution over parameters, which in effect, can water down the potential gains in predictive accuracy that a model with more parameters can achieve [@lee2014bayesian]. For this analysis, we treated age continuously.
